The World Bank, Women, Business and the Law

### 2024-02-21 outputs/paper/paper.qmd
fixed graph

Despite the many decades' worth of progress in gender equality throughout the world, one of the enduring types of gender discrimination worldwide is in the eyes of the Law. Though much progress has been made in western, high-income countries, there remains vast swaths of the world where women may not vote, work, or participate in the same activities as men do, due to the presence of gendered laws. To trace the progression of economic opportunities of women in the world's 190 countries, the World Bank constructed their "Women, Business, and the Law" (WBL) dataset [@world_bank], which Hyland et. al make of primary interest in their paper "Gendered Laws and Women in the Workforce"[@hylandetal]. In Sections 2, we analyze and replicate some of the key graphs and insights from their paper. Building on their work in Sections 4, we call into question the surface level ratings attributed by the World Bank by cross-referencing with the on-ground realities of gender-based discrimination using data from Statistics Canada. The concluding section outlines the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.

@@

  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
  geom_col() +
  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
  theme_minimal() +
  ylim (NA, 100) +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))

@@

  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
  geom_col() +
  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
  theme_minimal() +
  ylim (NA, 100) +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))

@@

\newpage

>>>>

Despite the many decades' worth of progress in gender equality throughout the world, one of the enduring types of gender discrimination worldwide is in the eyes of the Law. Though much progress has been made in western, high-income countries, there remains vast swaths of the world where women may not vote, work, or participate in the same activities as men do, due to the presence of gendered laws. To trace the progression of economic opportunities of women in the world's 190 countries, the World Bank constructed their "Women, Business, and the Law" (WBL) dataset [@world_bank], which Hyland et. al make of primary interest in their paper "Gendered Laws and Women in the Workforce"[@hylandetal]. In Section 2, we analyze and replicate some of the key graphs and insights from their paper. Building on their work in Sections 3, we call into question the surface level ratings attributed by the World Bank by cross-referencing with the on-ground realities of gender-based discrimination using data from Statistics Canada. The estimand we wish to replicate are the individual indicators used by WBL, but in a Canadian context that is backed by empirical research. The concluding section outlines the importance that accurate data have in being able to fully understand the complex picture of women and the law throughout the world.

@@

  ggplot(data = region_avg_index, mapping = aes(x = Region, y = avg_index)) +
  geom_col() +
  geom_text(aes(label = sprintf("%.2f", avg_index)), vjust = -0.5, color = "black") +  # Add this line for data labels
  labs(title="Average WBL Index by World Region, 2019", x="Region", y="WBL Index") +
  theme_minimal() +
  ylim (NA, 100) +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))

@@

## Replication of Figure 1B: Indicator variables averages 2019

# isolate indicators 
average_indicator <- wb_data |> 
                        rename(
                          Mobility = GR1_mobility, 
                          Workplace = GR2_workplace, 
                          Pay = GR3_pay, 
                          Marriage = GR4_marriage, 
                          Parenthood = GR5_parenthood, 
                          Entrepreneurship = GR6_entrep, 
                          Assets = GR7_assets, 
                          Pension = GR8_pension
                        ) |>
                        filter(`WBL Report Year` == 2019) |>
                        select(Mobility, 
                               Workplace, 
                               Pay, 
                               Marriage, 
                               Parenthood, 
                               Entrepreneurship, 
                               Assets, 
                               Pension
                               )

#average them out
avg_values <- colMeans(average_indicator)

#put them in a table 
new_table <- data.frame(Column_Names = names(avg_values), Average_Values = avg_values)
# Plot in ggplot2

  ggplot(data = new_table, mapping = aes(x = Column_Names, y = Average_Values)) +
  geom_col() +
  labs(title="Indicator Averages, 2019", x="Indicator", y="WBL Indicator Score") +
  theme_minimal() +
  ylim (NA, 100) +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text (angle = 45, vjust = 1, hjust=1))

@@

\newpage
Replication DOI Link: https://doi.org/10.48152/ssrp-qeab-5493
